Supplemental Materials, Methods and Results

### 2021-06-02 supp.Rmd
Added some more details on vs model, fixed up figS2

```{r Figcolsig, fig.align = 'center', fig.cap = "Additional examples of chameleon color signals and behavior in response to different social contexts. (A) Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male. (B) Two examples of males engaged in horn locking. Males in the bottom plane are evenly matched at this stage and both are in full display colour. In the top image, the male on the left is starting to change to a subordinate color. (C) Conspicuous lemon-yellow male courtship color. In this case, the female is using a contrasting stress color consistent with rejection."}
figs1 <- image_read("./output/figures/FigS1.png")
figs1

@@

```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Hawaii (A, C) and Kenya (B, D). In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines."}
figs2 <- image_read("./output/figures/FigS2.png")
figs2

@@

```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials. We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male."}
figs3 <- image_read("./output/figures/FigS3.jpg")
figs3

@@

## Color measurements, habitat background and irradiance

During all behavior trials, we measured the color of each body region (Fig. \@ref(fig:Figcolsig5)) once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal. Patches were measured in random sequence to avoid any bias of order. When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source. We used a custom-made probe-holder which we placed in contact with the animal at a standard angle (45&deg; @RN2) and distance (5 mm), thereby excluding ambient light. All measurements were relative to a dark, and a certified 99% white reflectance standard. When chameleons started to change color in response to our handling we stopped measuring color and allowed them to continue a social trial before continuing to measure spectral reflectance. In the case of a predator trial we would once again present them with the predator model until they returned to their anti-predator display. This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.

We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signalling environment. This included 1.) the upper surface of leaves; 2.) the underside of leaves; and 3.) branches. We also measured the habitat light environment by recording irradiance in the chameleons environment.

@@

@@

Chromatic and achromatic contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN4 in combination with the package `clubSandwich` @RN5. For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems). While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons. As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*). In all cases, our results and conclusions were not impacted (see supplementary code). Deriving multiple contrast for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other. To ensure that non-independence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction. Across the different contexts (i.e., social contexts, predation, and background environments), chromatic and achromatic contrasts were analyzed in separate models as response variables. For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects. For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects. Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were observed on a native Kenyan background or their current introduced habitat background in Hawaii. Relevant model interactions were first tested using Wald tests. If found to be non-significant (p > 0.1) we dropped the interaction and fit a main effects model. We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.

>>>>

```{r Figcolsig, fig.align = 'center', fig.cap = "Additional examples of chameleon color signals and behavior in response to different social contexts. (A) Two different males in conspicuous display color adopted at the start of a contest and retained by the dominant male. (B) Two examples of males engaged in horn locking. Males in the bottom plane are evenly matched at this stage and both are in full display colour. In the top image, the male on the left is starting to change to a subordinate color. (C) Conspicuous lemon-yellow male courtship color. In this case, the female is in a neutral state. When females reject males they develop a heavily contrasting pattern."}
figs1 <- image_read("./output/figures/FigS1.png")
figs1

@@

```{r Figcolsig2, fig.align = 'center', fig.cap = "Representative photos of habitat in Kenya (A, B) and Hawaii (C, D). In both sites, chameleons favored well-vegetated trees and bushes, often with lots of vines"}
figs2 <- image_read("./output/figures/FigS2.jpg")
figs2

@@

```{r Figcolsig3, fig.align = 'center', fig.cap = "The experimental setup for behavioral trials. We constructed a platform of three connected horizontal perches in a triangle, thereby always allowing an individual to escape a rival or if a female, a courting male. The male on the left won this contest and is in full display color while the male on the right has lost this contest and turned brown (subordinate)."}
figs3 <- image_read("./output/figures/FigS3.jpg")
figs3

@@

## Spectroradiometry, habitat background and irradiance

During all behavior trials, we measured the color of each body region (Fig. \@ref(fig:Figcolsig5)) once only, because chameleons change color rapidly and we had to reduce the handling stress to the animal. Patches were measured in random sequence to avoid any bias of order. When a chameleon had completely changed color (to a human observer) and was also performing typical behavior associated with their display (e.g., head shakes, swaying, gaping, rapid approach or flee) we lightly restrained it with one hand, without removing it from its perch, and measured spectral reflectance using a 1.2 m bifurcated probe connected to an Ocean Optics USB-2000 optic spectrophotometer and a PX2 light source. We used an Ocean Optics RPH-1 probe-holder which we placed in contact with the animal at a standard angle (45&deg; @RN2) and distance (5 mm), thereby excluding ambient light. All measurements were relative to a dark, and a certified 99% white reflectance standard. When chameleons started to change color in response to our handling we stopped measuring color and allowed them to continue a social trial before continuing to measure spectral reflectance. In the case of a predator trial we would once again present them with the predator model until they returned to their anti-predator display. This system worked effectively because chameleons readily responded to social and anti-predator stimuli and quickly returned to natural behavior following handling.

We quantified the background against which chameleons signal by measuring the spectral reflectance of vegetation in their signalling environment. This included 1.) the upper surface of leaves; 2.) the underside of leaves; and 3.) stems We also measured the habitat light environment by recording irradiance in the chameleons environment. Measurements were taken parallel to the ground (side-welling irradiance) to mimic signaling conditions for chameleons—light illuminating the chameleon’s flank is more likely to play a role in signaling @RN8.

@@

We first smoothed all spectral reflectance files by a factor of 0.16 using the plotsmooth function in Pavo (vers 1.0). We then lifted negative reflectance spectra by the factor they were negative by, thereby maintaining the shape of the curve. We also took the median reflectance for all leaves (top, bottom) and stems (2 measures/stem) for each population (Hawaii, Kenya), and for irradiance.

We used visual modelling to determine how conspicuous a chameleon would appear to both a conspecific receiver and a predator (bird and snake) using the Vorobyev-Osario (V-O) receptor noise model @RN9. The V-O model calculates perceptual distance (∆S) between two color patches in units of just noticeable differences (JNDs). A difference of at least 1 JND is thought to enable a receiver to discriminate two colors @RN10. We calculated JNDs for chromatic- and achromatic contrast for all body regions combined (overall conspicuousness) and for individual body regions in order to compare conspicuousness as a function of population (Hawaii, Kenya) or predator (bird, snake) in relation to background (vegetation). The receptor noise model incorporates the median reflectance of background color, irradiance, chameleon spectral reflectance and visual pigment absorbance functions for the Jackson’s chameleon (Loew unpubl. data), birds, and the garter snake @RN11.

@@

Chromatic and achromatic contrast from visual models, measured in just noticeable differences (JNDs), were analyzed using linear mixed effects models in R using the `lme4` package @RN4 in combination with the package `clubSandwich` @RN5. For all individuals, we measured signals for multiple body regions, each of which were contrasted between multiple background measurements (i.e., top and bottoms of two leaves and two stems). While we present sensitivity analyses for all body regions (see below - *Additional Supplemental Results*), we were mainly interested in overall signal differences displayed by chameleons. As such, for our main analyses we averaged estimates (marginalized) across these body regions and background environments given that in many cases body regions showed remarkably similar effects (see *Additional Supplemental Results*). In all cases, our results and conclusions were not impacted (see supplementary code). Deriving multiple contrasts for each individual body region and background, however, resulted in a substantial number of comparisons, each of which were not completely independent of each other. To ensure that non-independence of data did not impact our inferences we used a Robust Variance Estimator (RVE) (using our final models to correct standard errors) and a Satterthwaite degrees of freedom correction. Across the different contexts (i.e., social contexts, predation, and background environments), chromatic and achromatic contrasts were analyzed in separate models as response variables. For analyses involving social context, we included z-transformed snout-vent length (SVL), population of origin, and their interaction as fixed effects. For models involving predatory contexts and visual systems, we included the predator type (i.e, snake or bird), population of origin and their interaction as fixed effects. Finally, to test local adaptation, we used only chameleons from Hawaii and tested the extent to which chromatic and achromatic contrast varied depending on whether Hawaiian signals were observed on a native Kenyan background or their current introduced habitat background in Hawaii. Relevant model interactions were first tested using Wald tests. If found to be non-significant (p > 0.1) we dropped the interaction and fitted a main effects model. We then tested the significance of the main effect of 'Population', 'Predator' or 'Background' (depending on the model) using Wald tests with robust variance estimators and a Satterthwaite degrees of freedom correction.
